Structured matrix

### 2019-07-04 doc/fast_ssa/paper.tex
review the property of Fourier transform

\usepackage{xy}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{url}
\usepackage{rotating}
\usepackage{lscape}
\usepackage{color}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amssymb}      
\usepackage{algpseudocode}
\usepackage{booktabs}
 \usepackage{tikz}
 \usetikzlibrary{matrix}

@@

@@

Suppose we need to compute the product between Toeplitz matrix ${\bf T}$ and a vector ${\bf v}$
\begin{equation}
{\bf r} = {\bf T}{\bf v} \,
\end{equation} 
where the length of ${\bf v}$ is $K$.
There are two way to do it efficiently via FFT, and these two ways corresponding to the two embedding methods. The first step is to padding vector ${\bf v}$ with $L-1$ $0$s get a new vector $\hat{\bf v}$. Following above example, $K=L=5$ and ${\bf v} = \begin{bmatrix} v_1 & v_2 & v_3 & v_4 & v_5  \end{bmatrix}^T$, and the padded vector $\hat{\bf v} = \begin{bmatrix} v_1 & v_2 & v_3 & v_4 & v_5 & 0 & 0 & 0 & 0 \end{bmatrix}^T$, The product between ${\bf C}_1$ and $\hat{\bf v}$ are
\begin{equation}
\hat{\bf r}_1 = {\bf C}_1 \hat{\bf v} = \begin{bmatrix}
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
\end{bmatrix}
\begin{bmatrix}
v_1 \\
v_2 \\
v_3 \\
v_4 \\
v_5 \\
0 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\end{equation}
where the last $L$ elements of $\hat{\bf r}_1$ equal to ${\bf r}$. On the other hand, we can take advantage of the circulate matrix ${\bf C}_2$, we have
\begin{equation}
\hat{\bf r}_2 = {\bf C}_2 \hat{\bf v} = \begin{bmatrix}
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
\end{bmatrix}
\begin{bmatrix}
v_1 \\
v_2 \\
v_3 \\
v_4 \\
v_5 \\
0 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\end{equation}
So the first $L$ elements of $\hat{\bf r}_2$ equal to ${\bf r}$. We summarize the efficient computation of the product between Toeplitz matrix and a vector in Algorithm 1 and Algorithm 2
\begin{algorithm}
      \SetAlgoLined
      \Fn{ ${\bf v}= ttv({\bf c}, {\bf v}) $}{
      length of ${\bf c}$: $N = length({\bf c})$   \\
      compute $L = \floor*{\frac{N}{2}}+1$, $K = N+1 - L$   \\
      padding $L-1$ $0$s to the end of ${\bf v}$ to get a new vector $\hat{\bf v}$ with length of $N$\\
      ${\bf r}_1 = {\bf F}_N^{-1} \left(   {\bf F}_N {\bf c} \circ {\bf F}_N  \hat{\bf v}\right) $ \\
      take the last $L$ elements from ${\bf r}_1$ 
      }
      \caption{Teoplitz-Times-Vector}
\end{algorithm}

Another method is summarized in Algorithm 2 as 
\begin{algorithm}
      \SetAlgoLined
      \Fn{ ${\bf v}= ttv({\bf c}, {\bf v}) $}{
      length of ${\bf c}$: $N = length({\bf c})$   \\
      compute $L = \floor*{\frac{N}{2}}+1$, $K = N+1 - L$   \\
      padding $L-1$ $0$s to the end of ${\bf v}$ to get a new vector $\hat{\bf v}$ with length of $N$\\
      move the first $K-1$ element of ${\bf c}$ to the end, get a new vector $\hat{\bf c}$
      ${\bf r}_2 = {\bf F}_N^{-1} \left(   {\bf F}_N \hat{\bf c} \circ {\bf F}_N  \hat{\bf v}\right) $ \\
      take the first $L$ elements from ${\bf r}_2$ 
      }
      \caption{Teoplitz-Times-Vector}
\end{algorithm}

@@

One can convert a Hankel matrix to a Toeplitz matrix by reversing its columns. The columns reversing process can be represented by right multiplying a $K \times k$ matrix ${\bf R}$, whose elements are $1$ on the main anti-diagonals and 0 elsewhere. Follow the toy example, we can get
\begin{equation}
{\bf T}= \begin{bmatrix}
c_5 & c_4 & c_3 & c_2 & c_1 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 \\ 
\end{bmatrix} = {\bf HR} = 
\begin{bmatrix}
c_1 & c_2 & c_3 & c_4 & c_5 \\ 
c_2 & c_3 & c_4 & c_5 & c_6 \\ 
c_3 & c_4 & c_5 & c_6 & c_7 \\ 
c_4 & c_5 & c_6 & c_7 & c_8 \\ 
c_5 & c_6 & c_7 & c_8 & c_9 \\ 
\end{bmatrix}
\begin{bmatrix}
\quad & \quad &  \quad &  \quad & 1 \\
\quad & \quad &  \quad &  1 & \quad \\
\quad & \quad &  1 &  \quad & \quad \\
\quad & 1 &  \quad &  \quad & \quad \\
1 & \quad &  \quad &  \quad & \quad \\
\end{bmatrix}
\end{equation}

Before we discuss about the multiplication between Hankel matrix and a vector, we first look at the property of the column reversing matrix ${\bf R}$, we have 
\begin{equation}
{\bf R} = {\bf R}^T = {\bf R}^{-1}
\end{equation}
So the product between Hankel matrix ${\bf H}$ and a vector ${\bf v}$ is given as 
\begin{equation}
{\bf r} = {\bf Hv} = {\bf HRR}^{-1}{\bf v} = {\bf HRRv} = ({\bf HR})({\bf Rv}) = {\bf T} \hat {\bf v}
\end{equation}
where $\bf{v}$ is the reversed version of the vector ${\bf v}$.

>>>>

\usepackage{xy}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{url}
\usepackage{rotating}
\usepackage{lscape}
\usepackage{color}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amssymb}      
\usepackage{algpseudocode}
\usepackage{booktabs}
 \usepackage{tikz}
 \usetikzlibrary{matrix}
 \DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n}

@@

\subsection{Properties of Fourier transform}

In this section, we review the properties of discrete Fourier transform ({\bf DFT}), most of the properties are derived in continuous domain as it is much easier for the derivation, but the discrete version are verified by $Julia$ code.

We define the Fourier transform as
\begin{equation}
X(\omega) = \int_{-\infty}^{\infty} x(t) e^{-i\omega t} dt
\end{equation}
where $\omega$ is radian frequency and $t$ is time. The above operation can be simplified as 
\begin{equation}
{\mathbfcal F}[x(t)] \rightarrow X(\omega)
\end{equation}

\subsubsection{1. time shift}
\begin{equation}
{\mathbfcal F}[x(t-t_0)] = X(\omega)e^{i \omega t_0}
\end{equation}

{\bf Proof}:
$$
{\mathbfcal F}[x(t-t_0)]  = \int_{-\infty}^{\infty} x(t - t_0) e^{-i\omega t} dt  
$$
we set $t^{\prime} = t - t_0$ and exchange the variables, we can get
$$
{\mathbfcal F}[x(t-t_0)]  = \int_{-\infty}^{\infty} x(t^{\prime}) e^{-i\omega (t^{\prime} + t_0)} d t^{\prime}  $$
$$
{\mathbfcal F}[x(t-t_0)] = e^{i \omega t_0} \int_{-\infty}^{\infty} x(t^{\prime}) e^{-i\omega t^{\prime}} d t^{\prime} = X(\omega) e^{i \omega t_0}
$$

\subsubsection{2. frequency modulation}
\begin{equation}
{\mathbfcal F}[x(t)e^{i \omega_0 t}] = X(\omega-\omega_0)
\end{equation}
{\bf Proof}:
$$
{\mathbfcal F}[x(t)e^{i \omega_0 t}]  = \int_{-\infty}^{\infty} x(t)e^{i \omega_0 t} e^{-i\omega t} dt  \\
=  \int_{-\infty}^{\infty} x(t) e^{-i (\omega-\omega_0) t} dt  =  X(\omega-\omega_0)
$$
\textcolor{red}{Note: this is the property we forget very often.}

\subsubsection{3. time reversal}
\begin{equation}
{\mathbfcal F}[x(-t)] = X(-\omega)
\end{equation}
{\bf Proof}:
$$ {\mathbfcal F}[x(-t)]  = \int_{-\infty}^{\infty} x(-t)e^{-i \omega t} dt  $$
Set $t^{\prime} = -t$ and change the variables we can get
$$ {\mathbfcal F}[x(-t)]  = \int_{\infty}^{-\infty} x(t^{\prime}) e^{-i \omega (-t^{\prime})} d (-t^{\prime}) \\
= \int_{-\infty}^{\infty} x(t^{\prime}) e^{-i (-\omega) t^{\prime}} d t^{\prime} \\
= X(-\omega)$$
Based on this property, we can get two extensions, if the time function is and even function, that is
$$x(t) = x(-t) \rightarrow X(\omega) = X(-\omega)$$
if the time function is odd function, the Fourier transform of this function is also an odd function
$$x(t) = -x(-t) \rightarrow X(\omega) = - X(-\omega)$$

\subsubsection{4. time and frequency stretching}
\begin{equation}
{\mathbfcal F}[x(at)] = \frac{1}{a}X(\frac{\omega}{a})
\end{equation}
where $a$ is positive number

{\bf Proof}:
$$ {\mathbfcal F}[x(at)]  = \int_{-\infty}^{\infty} x(at)e^{-i \omega t} dt  $$
Set $t^{\prime} = a t$, By change the variable, we can get
$$ {\mathbfcal F}[x(at)]  = \int_{-\infty}^{\infty} x(t^{\prime})e^{-i \omega \frac{t^{\prime}}{a}} d\frac{t^{\prime}}{a} = \frac{1}{a}  \int_{-\infty}^{\infty} x(t^{\prime}) e^{-i \frac{\omega}{a} t^{\prime}} d t^{\prime} = \frac{1}{a}X{(\frac{w}{a})} $$
\textcolor{red}{Stretching in one domain means squeezing in another domain and vis versa.}

\subsubsection{5. Complex conjugate}
\begin{equation}
{\mathbfcal F}[x^*(t)] = X^*(- \omega)
\end{equation}
where the super-script $^*$ represent complex conjugate.

{\bf Proof}:
$$\int_{-\infty}^{\infty}  x^*(t) e^{-i \omega t} dt = \int_{-\infty}^{\infty} \left( x(t) e^{i \omega t}\right)^* dt = \int_{-\infty}^{\infty} \left(x(t) e^{-i (-\omega) t}\right)^* dt = \left( \int_{-\infty}^{\infty} x(t) e^{-i (-\omega) t} dt\right)^* = X^*(-\omega)  $$

@@

Suppose we need to compute the product between Toeplitz matrix ${\bf T}$ and a vector ${\bf v}$
\begin{equation}
{\bf r} = {\bf T}{\bf v} \,
\end{equation} 
where the length of ${\bf v}$ is $K$.
There are two way to do it efficiently via FFT, and these two ways corresponding to the two embedding methods. The first step is to padding vector ${\bf v}$ with $L-1$ $0$s get a new vector $\hat{\bf v}$. Following above example, $K=L=5$ and ${\bf v} = \begin{bmatrix} v_1 & v_2 & v_3 & v_4 & v_5  \end{bmatrix}^T$, and the padded vector $\hat{\bf v} = \begin{bmatrix} v_1 & v_2 & v_3 & v_4 & v_5 & 0 & 0 & 0 & 0 \end{bmatrix}^T$, The product between ${\bf C}_1$ and $\hat{\bf v}$ are
\begin{equation}
\hat{\bf r}_1 = {\bf C}_1 \hat{\bf v} = \begin{bmatrix}
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
\end{bmatrix}
\begin{bmatrix}
v_1 \\
v_2 \\
v_3 \\
v_4 \\
v_5 \\
0 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\end{equation}
where the last $L$ elements of $\hat{\bf r}_1$ equal to ${\bf r}$. On the other hand, we can take advantage of the circulate matrix ${\bf C}_2$, we have
\begin{equation}
\hat{\bf r}_2 = {\bf C}_2 \hat{\bf v} = \begin{bmatrix}
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
\end{bmatrix}
\begin{bmatrix}
v_1 \\
v_2 \\
v_3 \\
v_4 \\
v_5 \\
0 \\
0 \\
0 \\
0 \\
\end{bmatrix}
\end{equation}
So the first $L$ elements of $\hat{\bf r}_2$ equal to ${\bf r}$. We summarize the efficient computation of the product between Toeplitz matrix and a vector in Algorithm 1 and Algorithm 2
\begin{algorithm}
      \SetAlgoLined
      \Fn{ ${\bf r}= ttv({\bf c}, {\bf v}) $}{
      length of ${\bf c}$: $N = length({\bf c})$   \\
      compute $L = \floor*{\frac{N}{2}}+1$, $K = N+1 - L$   \\
      padding $L-1$ $0$s to the end of ${\bf v}$ to get a new vector $\hat{\bf v}$ with length of $N$\\
      ${\bf r}_1 = {\bf F}_N^{-1} \left(   {\bf F}_N {\bf c} \circ {\bf F}_N  \hat{\bf v}\right) $ \\
      take the last $L$ elements from ${\bf r}_1$ 
      }
      \caption{Teoplitz-Times-Vector}
\end{algorithm}

Another method is summarized in Algorithm 2 as 
\begin{algorithm}
      \SetAlgoLined
      \Fn{ ${\bf r}= ttv({\bf c}, {\bf v}) $}{
      length of ${\bf c}$: $N = length({\bf c})$   \\
      compute $L = \floor*{\frac{N}{2}}+1$, $K = N+1 - L$   \\
      padding $L-1$ $0$s to the end of ${\bf v}$ to get a new vector $\hat{\bf v}$ with length of $N$\\
      move the first $K-1$ element of ${\bf c}$ to the end, get a new vector $\hat{\bf c}$
      ${\bf r}_2 = {\bf F}_N^{-1} \left(   {\bf F}_N \hat{\bf c} \circ {\bf F}_N  \hat{\bf v}\right) $ \\
      take the first $L$ elements from ${\bf r}_2$ 
      }
      \caption{Teoplitz-Times-Vector}
\end{algorithm}

@@

One can convert a Hankel matrix to a Toeplitz matrix by reversing its columns. The columns reversing process can be represented by right multiplying a $K \times K$ matrix ${\bf R}$, whose elements are $1$ on the main anti-diagonals and 0 elsewhere. Follow the toy example, we can get
\begin{equation}
{\bf T}= \begin{bmatrix}
c_5 & c_4 & c_3 & c_2 & c_1 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 \\ 
\end{bmatrix} = {\bf HR} = 
\begin{bmatrix}
c_1 & c_2 & c_3 & c_4 & c_5 \\ 
c_2 & c_3 & c_4 & c_5 & c_6 \\ 
c_3 & c_4 & c_5 & c_6 & c_7 \\ 
c_4 & c_5 & c_6 & c_7 & c_8 \\ 
c_5 & c_6 & c_7 & c_8 & c_9 \\ 
\end{bmatrix}
\begin{bmatrix}
\quad & \quad &  \quad &  \quad & 1 \\
\quad & \quad &  \quad &  1 & \quad \\
\quad & \quad &  1 &  \quad & \quad \\
\quad & 1 &  \quad &  \quad & \quad \\
1 & \quad &  \quad &  \quad & \quad \\
\end{bmatrix}
\end{equation}

Before we discuss about the multiplication between Hankel matrix and a vector, we first look at the property of the column reversing matrix ${\bf R}$, we have 
\begin{equation}
{\bf R} = {\bf R}^T = {\bf R}^{-1}
\end{equation}
So the product between Hankel matrix ${\bf H}$ and a vector ${\bf v}$ is given as 
\begin{equation}
{\bf r} = {\bf Hv} = {\bf HRR}^{-1}{\bf v} = {\bf HRRv} = ({\bf HR})({\bf Rv}) = {\bf T} \hat {\bf v}
\end{equation}
where $\bf{v}$ is the reversed version of the vector ${\bf v}$. From above derivations, we can see that the multiplication between a Hankel matrix and a vector is converted to the multiplication between Teoplitz matrix and a vector. So the algorithm for the multiplication between Hankel matrix and a vector is summarized in Algorithm 3 as
\begin{algorithm}
      \SetAlgoLined
      \Fn{ ${\bf r}= htv({\bf c}, {\bf v}) $}{
      length of ${\bf c}$: $N = length({\bf c})$   \\
      compute $L = \floor*{\frac{N}{2}}+1$, $K = N+1 - L$   \\
      reverse the order of the elements of ${\bf v}$: $\hat{\bf v} = reverse({\bf v})$ \\
      call the function Toeplitz-times-vector: ${\bf r} = ttv({\bf c}, \hat{\bf v})$ \\
      \Return {\bf r}}
      \caption{Hankel-Times-Vector}
\end{algorithm}
